---
title: "PSTAT 131 Homework 2"
author: "Brett Goldman"
date: "10/3/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup
```{r}
library(tidyverse)
library(tidymodels)
```

```{r}
abalone<-read_csv(file="~/Documents/School/PSTAT 131/homework-2/data/abalone.csv")
```

# Question 1:

```{r}
abalone$age<-abalone$rings+5
hist(abalone$age, main="Histogram of Abalone Age", xlab="Age")
```
Looking at the histogram of age, most abalone in this dataset are between 12-15 years old. There are some very young and very old abelone as well, but it's mostly in that 12-15 range.

# Question 2:

```{r}
# 80/20 split
set.seed(3005)
abalone_split<-initial_split(abalone, prop=0.8, strata=age)
abalone_train<-training(abalone_split)
abalone_test<-testing(abalone_split)
```


# Question 3:
You shouldn't use rings to predict age because age is just rings+5. They have the same distribution just shifted over 5.
```{r}
simple_abalone_recipe<-
  recipe(age~type+longest_shell+diameter+height+whole_weight+
           shucked_weight+viscera_weight+shell_weight, data=abalone_train) %>% # recipe
  step_dummy(all_nominal_predictors()) 
```
%>% # dummy

```{r}
abalone_recipe<-simple_abalone_recipe %>%
  step_interact(terms = ~starts_with("type"):shucked_weight + longest_shell:diameter + shucked_weight:shell_weight)
```

```{r}
abalone_recipe<-abalone_recipe %>% 
  step_center(starts_with("type"), longest_shell, diameter, height, whole_weight,
           shucked_weight, viscera_weight, shell_weight)
```

```{r}
abalone_recipe<-abalone_recipe %>% 
  step_scale(starts_with("type"), longest_shell, diameter, height, whole_weight,
             shucked_weight, viscera_weight, shell_weight) # scale
```

# Question 4:
```{r}
abalone_lm_model<-linear_reg() %>% 
  set_engine("lm")
```

# Question 5:
```{r}
abalone_lm_wflow<-workflow() %>% 
  add_model(abalone_lm_model) %>% 
  add_recipe(abalone_recipe)
```

# Question 6:
```{r}
abalone_lm_fit<-fit(abalone_lm_wflow, abalone_train)
Q6Abalone<-data.frame(type="F", longest_shell=.5, diameter=.1, height=.3, 
                      whole_weight=4, shucked_weight=1, viscera_weight=2, shell_weight=1)
```

```{r}
(abalonepredict<-predict(abalone_lm_fit, new_data=Q6Abalone))
```

The age of a hypothetical female abalone with longest_shell = 0.50, diameter = 0.10, height = 0.30, whole_weight = 4, shucked_weight = 1, viscera_weight = 2, shell_weight = 1 is 17.92631.

# Question 7:
```{r}
library(yardstick)
abalone_train_res<-predict(abalone_lm_fit, new_data = abalone_train %>% select(-age))
abalone_train_res %>% 
  head()
```

```{r}
abalone_train_res<-bind_cols(abalone_train_res, abalone_train %>% select(age))
abalone_train_res %>% 
  head()
```
```{r}
rmse(abalone_train_res, truth=age, estimate=.pred)
```

```{r}
abalone_metrics<-metric_set(rmse, rsq, mae)
abalone_metrics(abalone_train_res, truth=age, estimate=.pred)
```

